---
title: "Zong-Nguyen"
author: "Maureen Cowan"
date: "2023-01-9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gene Ontology Analysis

### Bubble plot
```{r}
library(ggplot2)
library(dplyr)
library(data.table)


go <- read.csv("GO_up.csv")

go_interest <- go[c(2, 4, 5, 19, 21, 22, 29, 31, 33, 37, 45), ]

go_interest <- data.table(go_interest) #need data.table library for this

# Convert Gene ratio from character to numeric vector
frac <- go_interest$GeneRatio
frac <- sapply(frac, function(x) eval(parse(text=x)))


go_interest$Ratio <- frac
go_interest <- go_interest[order(-Count), ] #order function requires data.table


bubble_plot_V1 <- ggplot(data = go_interest, 
                      aes(x=Ratio, y=reorder(Description, Count), size = Count, fill = p.adjust)) + 
                      geom_point(alpha=0.7, shape = 21) +
                      scale_size(range = c(2.5, 8)) +
                      scale_fill_continuous(low = "blue", high = "red") +
                      theme_bw() +
                      xlab("Gene Ratio") + 
                      ylab("")

bubble_plot_V1

ggsave("bubble_plot_V1.tiff", dpi = 300)
```


### Bubble plot, V2
```{r}
go_interest <- go[c(20, 38, 2, 5, 12, 33, 107, 55, 45, 46), ]

go_interest <- data.table(go_interest) #need data.table library for this

# Convert Gene ratio from character to numeric vector
frac <- go_interest$GeneRatio
frac <- sapply(frac, function(x) eval(parse(text=x)))


go_interest$Ratio <- frac
go_interest <- go_interest[order(-Count), ] #order function requires data.table


bubble_plot_V2 <- ggplot(data = go_interest, 
                      aes(x=Ratio, y=reorder(Description, Count), size = Count, fill = p.adjust)) + 
                      geom_point(alpha=0.7, shape = 21) +
                      scale_size(range = c(2.5, 8)) +
                      scale_fill_continuous(low = "blue", high = "red") +
                      theme_bw() +
                      xlab("Gene Ratio") + 
                      ylab("")

bubble_plot_V2

ggsave("bubble_plot_V2.tiff", dpi = 300)
```

# Differential Gene Expression
### Enhanced Volcano
```{r}
#BiocManager::install("EnhancedVolcano") ## Install first session only

library(EnhancedVolcano)
results <- read.csv("sham_vs_toxo_DESeq2.txt", sep="", header = TRUE)

Volcano <- EnhancedVolcano(results,
    lab = results$gene_id,
    selectLab = c('Ubd', 'H2-Ab1', 'H2-Oa', 'Stat1', 'Tgtp1', 'Cxcl9', 'Gbp2b', 'Ciita', 'Il1rl1', 'Fcrls', 'Tgfb2'),
    x = 'log2FoldChange',
    y = 'pvalue',
    title = 'Sham vs. Toxo Macrophages',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    pointSize = 2.0,
    labSize = 7,
    col=c('grey', 'grey', 'grey', 'blue'),
    shape = 1,
    colAlpha = 0.5,
    legendLabSize = (8), # size of gene names
    xlim = c(-5, 8.5),
    widthConnectors = 0.75,
    drawConnectors = TRUE,
    legendPosition = 'none')

Volcano

#Il1rl1 is ST2. We have shown that during Toxo infection, it is downregulated (Katherine Still, PLOS Pathogens)

#options(ggrepel.max.overlaps = Inf)

ggsave("Volcano.tiff", dpi = 300)
```


# CountData: Heatmaps

```{r}
countData <- read.csv("count_TAM_sham_vs_toxo.txt", sep=",", header = TRUE, row.names= 1)

rld <- rlog(countData)

write.csv(assay(rld), "rld_counts.csv", sep = ",", row.names = 1)

```

## DESeqDataSetFromMatrix

To use DESeqDataSetFromMatrix, the user should provide the counts matrix, the information about the samples (the columns of the count matrix) as a DataFrame or data.frame, and the design formula.

### Setup
```{r}
#BiocManager::install("DESeq2") ## Install first session only
library(DESeq2)
```

### Define Arguments
```{r}
#colData: DF with at least one column. Rows correspond to columns of countData.
colData <- data.frame(condition = factor(c("Toxo", "Sham", "Toxo", "Sham", "Sham", "Toxo", "Toxo", "Toxo", "Sham", "Sham")))

rownames(colData) <- c("M_19116_S9", "M_19120_S2", "M_19121_S10", "M_19122_S3", "M_19184_S1", "M_19190_S8", "M_19193_S7", "M_19198_S6", "M_19510_S4", "M_19517_S5")


#cts: matrix
countData = as.matrix(read.csv("count_TAM_sham_vs_toxo.txt", sep=",", header = TRUE, row.names= 1)) ##variable name for countData matrix

  
#design: formula that expresses how the counts for each gene depend on the variables in colData. Use ~condition design since there are two groups being compared

design = ~condition #Design depends on condition as an explanatory variable

```


### Recreate the DESeq2 object
```{r}

dds <- DESeqDataSetFromMatrix(countData = (countData),
                              colData,
                              design)

## NOTE: Some of the values in the assay are not integers. These values have been normalized.

```

###  
```{r}
dds <- DESeq(dds)


res <- results(dds, contrast = c("condition","Sham", "Toxo")) ## define numerator and denominator. set the group you want on the left of a volcano plot first. 

#res <- results(dds)
```


## Import files
```{r}
#countData <- read.csv("count_TAM_sham_vs_toxo.txt", sep = ",")

```

